---
title: "Unsupervised Learning" 
subtitle: Machine Learning in Economics
author: 
  name: "Prof. Dr. Hüseyin Taştan"
  affiliation: "Yildiz Technical University"
date: "Spring 2021"
output:
  html_document: 
    number_sections: true
    theme: lumen
    highlight: haddock 
    # code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
    keep_md: false


---
<style type="text/css"> 
body{
  background-color: #FAFAFA;
  font-size: 18px;
  line-height: 1.8;
}
code.r{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE, results = 'asis', fig.show = 'asis')
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
``` 

<br/>
<br/>

# Principal Components

We will use `USArrests` data set which is a part of the base `R`.   
```{r} 
library(ISLR)
states <- row.names(USArrests) 
names(USArrests)
```


```{r}
apply(USArrests, 2, mean)
```

```{r}
apply(USArrests, 2, var) 
```

There are several packages to conduct PCA analysis (`prcomp`, `princomp`, etc.)
We standardize each variable using the option `scale=TRUE`
```{r}
pr.out <- prcomp(USArrests, scale=TRUE)
names(pr.out) 
```

The center and scale components correspond to the means and standard
deviations of the variables that were used for scaling prior to implementing
PCA.
```{r}
pr.out$center
```

```{r}
pr.out$scale
```

The `rotation` matrix provides the principal component loadings; each column
of `pr.out$rotation` contains the corresponding principal component
loading vector
```{r}
pr.out$rotation
```

Principle component score vectors: 
```{r}
dim(pr.out$x)  
```

We can plot the first two principal components as follows
```{r}
biplot(pr.out, scale=0, cex=0.6)
```

The `scale=0` argument to `biplot()` ensures that the arrows are scaled to 
represent the loadings; other values for scale give slightly different biplots
with different interpretations.

Notice that this figure is a mirror image of Figure 10.1. Recall that
the principal components are only unique up to a sign change, so we can
reproduce Figure 10.1 by making a few small changes
```{r}
pr.out$rotation <- -pr.out$rotation
pr.out$x <- -pr.out$x
biplot(pr.out, scale=0, cex=0.6)
```




Importance of principal components: 
```{r}
summary(pr.out)
```

We see that the first principal component explains 62.0% of the variance
in the data, the next principal component explains 24.7% of the variance,
and so forth.

```{r}
screeplot(pr.out, type="lines") 
```

```{r}
pr.out$sdev
pr.var <- pr.out$sdev^2
pr.var
pve <- pr.var/sum(pr.var)
pve
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')
plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

## Example: Body Data 

This data set contains a set of body measurements along with age, height, weight and gender on 507 individuals (247 men 260 women). We expect some of these variables to be highly correlated.  

```{r}
# body data from 
# Journal of Statistics Education Volume 11, Number 2 (2003),
# www.amstat.org/publications/jse/v11n2/datasets.heinz.html

load("../Data/body.RData") 
bodydata <- subset(body, select = -c(age, gender, gender_dummy)) 
str(bodydata)
```

Inspecting the correlation matrix we see that these features are in fact highly correlated. 
```{r}
library(ggcorrplot)
cormat <- round(cor(bodydata), 2)
ggcorrplot(cormat, hc.order = TRUE, type = "lower", outline.color = "white")
```

Find the principal components: 
```{r}
pr.out <- prcomp(bodydata, scale=TRUE)
summary(pr.out)  
# change the signs of factor loadings
pr.out$rotation <- -pr.out$rotation
pr.out$x <- -pr.out$x
biplot(pr.out, scale=0, cex=0.6)
```

```{r} 
pr.var <- pr.out$sdev^2 
pve <- pr.var/sum(pr.var)
pve
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')
plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

Approximately 85% of the total variation can be explained by the first 5 principal components. 


## Example: Regional Happiness Indicators

```{r}
# load the data set
load("../Data/lifedata2015.RData")
# exclude the first column and send it to prcomp() function 
pr_happiness <- prcomp(lifedata2015[,-1], scale=TRUE)
summary(pr_happiness)
```


```{r}
# change the signs of factor loadings
# and plot the biplot
pr_happiness$rotation <- -pr_happiness$rotation
pr_happiness$x <- -pr_happiness$x
# biplot
biplot(pr_happiness, scale=0, cex=0.6)
```

```{r} 
# compute proportion of variance explained
pr.var <- pr_happiness$sdev^2 
pve <- pr.var/sum(pr.var)
pve
```


```{r} 
# plot PVE
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')
```


```{r} 
# cumulative proportion of variance explained
plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

Approximately 80% of the variation in the data can be explained by the first 10 principal components. 

An alternative package:
```{r}
library(factoextra)
# PCA using factomineR
library(FactoMineR)
res_pca_lifedata <- PCA(lifedata2015[,-1],  graph = FALSE)
get_eig(res_pca_lifedata)
```


```{r}
# Scree plot
fviz_screeplot(res_pca_lifedata, addlabels = TRUE)
```



## Principal Components Regression (PCR)

Let's apply PCR approach to the regional happiness data. We want to predict provincial happiness levels using the first few principal components. 

We can use `pcr()` function in the `{pls}` package for that purpose. 

```{r}
library(pls)
set.seed(123)

pcr_fit <- pcr(happiness_level ~ . -province, 
               data = lifedata2015, 
               scale=TRUE, 
               validation="LOO") # use leave-one-out cross-validation
summary(pcr_fit)
```


```{r}
# how many principal components should we use? 
# RMSE as a function of the number of PC as computed using CV
validationplot(pcr_fit)
```

The minimum RMSE is achieved when we use 19 components. 
```{r}
pcr_fit <- pcr(happiness_level ~ . -province, 
               data = lifedata2015, 
               scale = TRUE, 
               ncomp = 19)
summary(pcr_fit)
```


# K-Means Clustering 

The function `kmeans()` performs K-means clustering in R. We begin with
a simple simulated example in which there truly are two clusters in the
data: the first 25 observations have a mean shift relative to the next 25
observations.

```{r}
# create a simulated data set
set.seed(2)
x <- matrix(rnorm(50*2), ncol=2)
x[1:25,1] <- x[1:25,1]+3
x[1:25,2] <- x[1:25,2]-4
```


```{r}
# perfom Kmeans clustering with 2 centers and 20 random starts
km.out <- kmeans(x, centers = 2, nstart = 20)
# print cluster information 
km.out$cluster
```


```{r}
# plot clusters
plot(x, col = (km.out$cluster+1), 
     main = "K-Means Clustering Results with K=2", 
     xlab = "", 
     ylab = "", 
     pch = 20, 
     cex = 2) 
```

In this example, we knew that there really were two clusters because
we generated the data. However, for real data, in general we do not know
the true number of clusters. We could instead have performed K-means
clustering on this example with K = 3.


```{r}
set.seed(4)
km.out <- kmeans(x, centers = 3, nstart = 20)
km.out
```


```{r}
# plot the results with k=3
plot(x, col = (km.out$cluster+1), 
     main = "K-Means Clustering Results with K=3", 
     xlab = "", 
     ylab = "", 
     pch = 20, 
     cex = 2)
```


When K = 3, K-means clustering splits up the two clusters.
It is strongly recommended to run K-means clustering with a large
value of `nstart`, such as 20 or 50, since otherwise an undesirable local
optimum may be obtained.

```{r}
# compute and print within sum of squares 
set.seed(3)
km.out <- kmeans(x,3,nstart=1)
km.out$tot.withinss
km.out <- kmeans(x,3,nstart=20)
km.out$tot.withinss
```

## Example: Bodydata 

```{r}
# the data set was loaded previously
km.out <- kmeans(bodydata, centers = 2, nstart = 20)
plot(bodydata[,1:3], col=(km.out$cluster), cex=2)
```

Plot clustering based on weight and height and compare to the actual gender factor: 
```{r} 
# Large blank circles are the resulting clusters
plot(bodydata[,22:23],col=(km.out$cluster),cex=2)
# put dots inside circles representing observed gender
# red = men, black = women
points(body[,23:24], col=c(1,2)[body$gender], pch=19, cex=1)
```
```{r} 
observed_class <- c(1,2)[body$gender]
km_cluster <- km.out$cluster
ratio <- sum(observed_class == km_cluster)/nrow(bodydata)
ratio
```

84% of observations seem to be clustered correctly. Note that if we know the clusters 
we do not need to apply K-Means clustering method. This was just for demonstration purposes. 
When we know the labels (like gender in the bodydata) there are better classification methods which we've discussed in our previous classes. 


## Example: Province-level well-being data 

In this example we will use Turkish province data set and its subset. 
```{r}
# load packages 
library(tidyverse)  # tidy data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization 
```


```{r} 
# load data 
load("../Data/lifedata2015.RData") 
# make province names rownames
lifedata2015 <- column_to_rownames(lifedata2015, var = "province")
```



```{r}
# use only happiness level and some selected variables 
happiness <- lifedata2015 %>% 
  select(happiness_level, life_exp, avr_daily_earnings, sat_rate_with_health_status)
head(happiness)
```

Display the correlation matrix of the happiness data: 
```{r}
library(ggcorrplot)
cormat <- round(cor(happiness), 2)
ggcorrplot(cormat, hc.order = TRUE, type = "lower", outline.color = "white")
```


```{r}
# visualize the distance function 
# blue: more similar; red: dissimilar 
# (look at the color legend, low value means more similar)
distance <- dist(scale(happiness))
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```
```{r}
# display few elements of distance matrix
as.matrix(distance)[1:6, 1:6]
```


Apply K-means clustering: 
```{r}
# find cluster 
happiness_scaled <- scale(happiness)
kmeans_happy <- kmeans(happiness_scaled, centers = 2, nstart=20)
# bivariate scatter plots
plot(happiness[,1:4],col=(kmeans_happy$cluster),cex=2)
```


```{r}
# cluster plot using principal components
# use factoextra::fviz_cluster() function to visualize clusters
# along with the first two principal components
fviz_cluster(kmeans_happy, data = happiness)
```

```{r}
# cluster plot using principal components
# without ellipses
fviz_cluster(kmeans_happy, geom = "point", ellipse = FALSE, data = happiness) +
  ggtitle("K-means cluster of provinces with K=2")
```



```{r}
# PCA using factomineR
library(FactoMineR)
res.pca <- PCA(happiness,  graph = FALSE)
get_eig(res.pca)
```


```{r}
# Scree plot
fviz_screeplot(res.pca, addlabels = TRUE, ylim = c(0, 50))
```


```{r}
# cluster plot using original data 
happiness %>%
  mutate(cluster  = kmeans_happy$cluster,
         province = row.names(happiness)) %>%
  ggplot(aes(happiness_level, avr_daily_earnings, color = factor(cluster), label = province)) +
  geom_text()
```


```{r}
# cluster plot using original data 
happiness %>%
  mutate(cluster  = kmeans_happy$cluster,
         province = row.names(happiness)) %>%
  ggplot(aes(happiness_level, life_exp, color = factor(cluster), label = province)) +
  geom_text()
```

```{r}
# cluster plot using original data 
happiness %>%
  mutate(cluster  = kmeans_happy$cluster,
         province = row.names(happiness)) %>%
  ggplot(aes(happiness_level, sat_rate_with_health_status, color = factor(cluster), label = province)) +
  geom_text()
```

```{r message=FALSE, warning=FALSE}
# compare and contrast different number of clusters
# plot different number of clusters 
k2 <- kmeans(happiness, centers = 2, nstart = 25)
k3 <- kmeans(happiness, centers = 3, nstart = 25)
k4 <- kmeans(happiness, centers = 4, nstart = 25)
k5 <- kmeans(happiness, centers = 5, nstart = 25)

# plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = happiness) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = happiness) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = happiness) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = happiness) + ggtitle("k = 5")

library(gridExtra)
grid.arrange(p1, p2, p3, p4, nrow = 2)
```

```{r}
# determine the optimal number of clusters 
set.seed(123)
# look for the elbow
fviz_nbclust(happiness, kmeans, method = "wss")
```



```{r}
# for optimal k=3 compute summary statistics 
happiness %>%
  mutate(Cluster = k3$cluster) %>%
  group_by(Cluster) %>%
  summarise_all("mean")
```

```{r}
# visualize K-means results with 4 clusters
fviz_cluster(k4, data = happiness,
             palette = c("#2E9FDF", "#28B463", "#E7B800", "#FC4E07"),
             ellipse.type = "euclid", # Concentration ellipse
             star.plot = TRUE,        # Add segments from centroids to items
             repel = TRUE,            # Avoid label overplotting 
             ggtheme = theme_minimal()
)
```
```{r}
# summary results for k = 4 clusters 
k4$tot.withinss # total sum of sq.
k4$centers      # centers (means) of clusers
k4$size         # number of obs. in clusters
```

```{r}
# for k=4 add cluster results to the data set 
# and compute summary statistics 
happiness_cl <- happiness %>%
  mutate(cluster_kmeans = k4$cluster) 

happiness_cl %>%
  group_by(cluster_kmeans) %>%
  summarise_all("mean")
```



# Hierarchical Clustering 

The `hclust()` function implements hierarchical clustering in R.


```{r}
# use previously created simulated data x (see kmeans clustering)
# clustering with different linkage functions 
hc.complete <- hclust(dist(x), method="complete")
hc.average <- hclust(dist(x), method="average")
hc.single <- hclust(dist(x), method="single")
hc.ward2 <- hclust(dist(x), method="ward.D2")
```

The option `"method="ward.D2"` performs Ward's minimum variance method. In this 
method the total within-cluster variance is minimized. It merges the pair of clusters 
with minimum between-cluster distance at each step. (This option corresponds 
to `cluster::agnes(..., method = "ward")` in the `{cluster}` package. 

Visualize them using base R plot function: 
```{r}
# plot them side-by-side using base R plot function 
# cex = label size
labsize = 0.8 
par(mfrow=c(2,2))
plot(hc.complete,main="Complete Linkage", xlab="", sub="", cex=labsize)
plot(hc.average, main="Average Linkage", xlab="", sub="", cex=labsize)
plot(hc.single, main="Single Linkage", xlab="", sub="", cex=labsize)
plot(hc.ward2, main="Ward Linkage", xlab="", sub="", cex=labsize)
```
 


```{r}
# cut dendrogram into groups of data 
# k = 2 groups
cutree(hc.complete, k= 2)
```

```{r}
cutree(hc.average, 2)
```


```{r}
cutree(hc.single, 2)
```


```{r}
# 4 clusters using single linkage
cutree(hc.single, 4)
```

To scale the variables before performing hierarchical clustering of the
observations, we use the `scale()` function:
```{r}
xsc <- scale(x)
plot(hclust(dist(xsc), method="complete"), main="Hierarchical Clustering with Scaled Features")
```

## Example: Province-level Well-being data 

```{r}
# load packages 
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization 
```


```{r}
# load data 
load("../Data/lifedata2015.RData") 
# make province names rownames
lifedata2015 <- column_to_rownames(lifedata2015, var = "province")
```


```{r}
# hierarchical clustering 
# using stats::hclust package
# scale variables before dist() function
# can also use pipes, %>%, see example at the end
hc_complete <- hclust(dist(scale(lifedata2015)), method="complete")
hc_average <- hclust(dist(scale(lifedata2015)), method="average")
hc_single <- hclust(dist(scale(lifedata2015)), method="single")
hc_ward <- hclust(dist(scale(lifedata2015)), method="ward.D2")
```


```{r}
labsize = 0.7
plot(hc_complete, main = "Complete Linkage", xlab="", sub="", cex=labsize)
```


```{r}
plot(hc_average, main = "Average Linkage", xlab="", sub="", cex=labsize)
```


```{r}
plot(hc_single, main = "Single Linkage", xlab="", sub="", cex=labsize)
```


```{r}
plot(hc_ward, main = "Ward's Method", xlab="", sub="", cex=labsize)
```



```{r}
# focus on a subset of variables 
# use only happiness level and some selected variables 
happiness <- lifedata2015 %>% 
  select(happiness_level, life_exp, avr_daily_earnings, sat_rate_with_health_status)
```
 


```{r}
# cluster using hclust
hc_complete <- hclust(dist(scale(happiness)), method="complete")
hc_average <- hclust(dist(scale(happiness)), method="average")
hc_single <- hclust(dist(scale(happiness)), method="single")
hc_ward <- hclust(dist(scale(happiness)), method="ward.D2")
```


```{r}
plot(hc_complete, main="Complete Linkage", xlab="", sub="", cex=labsize)
plot(hc_average, main="Average Linkage", xlab="", sub="", cex=labsize)
plot(hc_ward, main="Ward's method", xlab="", sub="", cex=labsize)
```


```{r}
# form clusters  
ncluster = 2
cutree(hc_complete, ncluster)
```


```{r}
# plot dendrogram
plot(hc_complete)
# draw rectangles around clusters
rect.hclust(hc_complete , k = ncluster, border = 2:6) 
```


```{r}
# colorize branches using dendextend library
library(dendextend)
avg_dend_obj <- as.dendrogram(hc_complete)
avg_col_dend <- color_branches(avg_dend_obj, h = 3)
plot(avg_col_dend)
```




```{r}
# Cut tree into  groups
hc_cluster2 <- cutree(hc_complete, k = 2)

# Number of members in each cluster
table(hc_cluster2)
```


```{r}
# visualize result in a scatter plot using factoextra package 
# 2 clusters
library(factoextra) 
fviz_cluster(list(data = happiness, cluster = hc_cluster2))
```

```{r}
# we can add cluster information into data 
happiness %>%
  mutate(cluster = hc_cluster2) %>%
  head
```


```{r}
# Cut tree into  3 groups
# use Ward method results
hc_cluster3 <- cutree(hc_ward, k = 3)

# Number of members in each cluster
table(hc_cluster3)
```


```{r}
# visualize dendrogram 
# use factoextra::hcut() function 
hres3 <- hcut(happiness, k = 3, stand = TRUE)
# Visualize
fviz_dend(hres3, rect = TRUE, cex = 0.5,
          k_colors = c("#F1C40F","#28B463", "#E67E22"))
# for html color codes visit https://htmlcolorcodes.com/ 
```



```{r}
# visualize result in a scatter plot using factoextra package 
# 3 clusters 
library(factoextra) 
fviz_cluster(list(data = happiness, cluster = hc_cluster3), 
             palette = c("#F1C40F", "#E67E22", "#28B463"), 
             repel = TRUE # avoid overplotting text labels
             )
```



```{r}
# visualize dendrogram 
# use factoextra::hcut() function 
# cut into 4 clusters
hres4 <- hcut(happiness, k = 4, stand = TRUE) # default is hc_method = "ward.D2"
# Visualize
fviz_dend(hres4, cex = 0.5,
          k_colors = c("#F1C40F", "#28B463", "#E67E22", "#3498DB"), 
          color_labels_by_k = TRUE,
          rect = TRUE,
          rect_border = c("#F1C40F","#28B463", "#E67E22", "#3498DB"), 
          rect_fill = TRUE
         )
```

```{r}
# horizontal dendrogram: 
# add horiz = TRUE option
fviz_dend(hres4, cex = 0.5, horiz = TRUE, 
          k_colors = c("#F1C40F", "#28B463", "#E67E22", "#3498DB"), 
          color_labels_by_k = TRUE,
          rect = TRUE,
          rect_border = c("#F1C40F","#28B463", "#E67E22", "#3498DB"), 
          rect_fill = TRUE
         )
```

```{r}
# circular dendrogram   
# add type = "circular" option
fviz_dend(hres4, cex = 0.5, type = "circular", 
          k_colors = c("#F1C40F", "#28B463", "#E67E22", "#3498DB") 
         )
``` 


```{r eval=FALSE, include=TRUE}
# Alternatively dendrogram may be cut inside the fviz_dend
fviz_dend(hc_ward,      # cluster results
          k = 4,        # desired number of clusters
          rect = TRUE,  # draw rectangles around clusters
          cex = 0.5     # label size (province names)
          )
```



```{r}
# 4 clusters using Ward's method 
# Cut tree into  4 groups
hc_cluster_ward4 <- cutree(hc_ward, k = 4)
library(factoextra) 
fviz_cluster(list(data = happiness, cluster = hc_cluster_ward4),
          palette = c("#F1C40F", "#3498DB", "#E67E22", "#28B463"), 
          repel = TRUE)
```

```{r}
# add hierarchical clustering result to the data set
happiness_cl <- happiness_cl %>% 
  mutate(cluster_hc = hres4$cluster)
# cluster means
happiness_cl %>% 
  select(-cluster_kmeans) %>% 
  group_by(cluster_hc) %>% 
  summarize_all(mean)
```



Illustration of using the pipe operator ` %>% `: 
```{r}
# using pipes
# dend <- 1:10 %>% dist %>% hclust %>% as.dendrogram 
# dend %>% plot()

dend <- happiness %>%               # take the data
  scale() %>%                       # standardize
  dist %>%                          # calculate a distance matrix, 
  hclust(method = "complete") %>%   # hierarchical clustering using  
  as.dendrogram                     # turn it into a dendrogram.

dend %>% plot()
```



<br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

