---
title: "Nonlinear Models" 
subtitle: Machine Learning in Economics
author: 
  name: "Prof. Dr. Hüseyin Taştan"
  affiliation: "Yildiz Technical University"
date: "Spring 2021"
output:
  html_document: 
    number_sections: true
    theme: readable
    highlight: haddock 
    # code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
    keep_md: false


---
<style type="text/css"> 
body{
  font-size: 12pt;
}
code.r{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE, results = 'asis', fig.show = 'asis')
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
``` 

<br/>
<br/>



# Polynomial Regression and Step Functions


We will use the `Wage` data set throughout the lab. 

```{r}
library(ISLR) 
head(Wage)
levels(Wage$education)
```

The following code chunk can be used to estimate the polynomial regression:
```{r}
fit <- lm(wage ~ poly(age, 4), data = Wage)
coef(summary(fit))
```

The `poly()` function uses orthogonal polynomials which are essentially linear combinations of the variables `age,age^2,age^3,age^4`. 

The function `poly()` can also be used directly without orthogonal polynomials using the raw option: 
```{r}
fit2 <- lm(wage ~ poly(age, 4, raw=T), data = Wage)
coef(summary(fit2))
```

While this does not affect the predictions, the coefficient estimates will change. 

Another option to estimate polynomial regression is to use
```{r}
fit2a <- lm(wage ~ age + I(age^2) + I(age^3) + I(age^4), data = Wage)
coef(fit2a)
```

or 
```{r}
fit2b <- lm(wage ~ cbind(age, age^2, age^3, age^4), data = Wage)
```

Create a grid of values for the variable `age` and then use the generic `predict()` function to obtain predictions along with standard errors. 
```{r}
# set range of age
agelims <- range(Wage$age)
age.grid <- seq(from=agelims[1], to=agelims[2])
# obtain predictions
preds <- predict(fit, newdata=list(age=age.grid), se=TRUE)
# confidence bands: 
se.bands <- cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
```

Plot to produce Figure 7.1 (left panel) 
```{r}
#par(mfrow=c(1,2), mar=c(4.5,4.5,1,1), oma=c(0,0,4,0))
plot(Wage$age, Wage$wage, xlim=agelims, cex=.5, col="darkgrey", 
     xlab = "Age", ylab = "Wage")
title("Degree-4 Polynomial", outer=T)
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, lwd=1, col="blue", lty=3)
```

How do we choose the degree of the polynomial? One way is to use the ANOVA approach in which we conduct sequential F tests to evaluate nested models. Consider a sequence of polynomials from degree 1 to 5. The degree 1 will be nested in the degree 2, the degree 2 will be nested in degree 3, etc. 

So, let's estimate sequential models of polynomials with increasing complexity and apply F tests using `anova()` function. 
```{r}
fit.1 <- lm(wage ~ age, data=Wage)
fit.2 <- lm(wage ~ poly(age,2),data=Wage)
fit.3 <- lm(wage ~ poly(age,3),data=Wage)
fit.4 <- lm(wage ~ poly(age,4),data=Wage)
fit.5 <- lm(wage ~ poly(age,5),data=Wage)
anova(fit.1, fit.2, fit.3, fit.4, fit.5) 
```

Compare linear Model 1 to quadratic Model-2: p-value is practically zero (< 2.2e-16) so we reject Model-1, i.e., linear model is not adequate. 

Compare the quadratic Model-2 to the cubic Model 3: p-value = 0.0017, we reject Model-2, so the quadratic model is not sufficient. 

Compare the cubic and degree-4 polynomials, Model 3 and Model 4: the p-value is approximately 5% while the degree-5 polynomial Model-5 seems unnecessary
because its p-value is 0.37. Hence, either a cubic or a quartic (4th degree) polynomial
appear to provide a reasonable fit to the data, but lower- or higher-order
models are not justified.
```{r}
coef(summary(fit.5))
```

**Exercise**: Try cross-validation to choose the polynomial degree. 

<br>

Now, let's fit a logistic regression with polynomial terms using `glm()`: 
```{r}
fit <- glm( I(wage>250) ~ poly(age, 4), data=Wage, family=binomial)
preds <- predict(fit, newdata=list(age=age.grid), se=T)

# obtain standard error bands 
pfit <- exp(preds$fit)/(1+exp(preds$fit))
se.bands.logit <- cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
se.bands <- exp(se.bands.logit)/(1+exp(se.bands.logit))
```

Plot the predicted probabilities and confidence bands (figure 7.1 right panel): 
```{r}
plot(Wage$age, I(Wage$wage>250), xlim=agelims, type="n", ylim=c(0,.2), 
     xlab = "Age", ylab = "Pr(wage>250)")
points(jitter(Wage$age), I((Wage$wage>250)/5),cex=.5,pch="|",col="darkgrey")
lines(age.grid,pfit,lwd=2, col="blue")
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)
```

In the graphs above, the `jitter()` function is used to prevent the same age observations from overlapping each other.

Put both graphs together as in Figure 7.1: 
```{r}
par(mfrow=c(1,2), mar=c(4.5,4.5,1,1), oma=c(0,0,4,0))
plot(Wage$age, Wage$wage, xlim=agelims, cex=.5, col="darkgrey", 
     xlab = "Age", ylab = "Wage")
title("Degree-4 Polynomial", outer=T)
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, lwd=1, col="blue", lty=3)

plot(Wage$age, I(Wage$wage>250), xlim=agelims, type="n", ylim=c(0,.2), 
     xlab = "Age", ylab = "Pr(wage>250)")
points(jitter(Wage$age), I((Wage$wage>250)/5),cex=.5,pch="|",col="darkgrey")
lines(age.grid,pfit,lwd=2, col="blue")
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)
```

The `cut()` function creates cutpoints in the observations, which are then used as predictors for the linear model to fit a step function.


```{r}
table(cut(Wage$age, 4))
```


```{r}
fit <- lm(wage ~ cut(age, 4), data = Wage)
coef(summary(fit))
```


# Splines 

In `R` we can use `splines` library to fit spline models. The `bs()` function is used to specify basis functions. The default is the cubic splines. For example, 

```{r}
library(splines)
fit <- lm(wage ~ bs(age, knots=c(25,40,60)), data=Wage)
pred <- predict(fit, newdata=list(age=age.grid), se=T)
plot(Wage$age, Wage$wage, col="gray", 
     xlab = "Age", ylab = "Wage")
lines(age.grid, pred$fit, lwd=2)
lines(age.grid, pred$fit+2*pred$se, lty="dashed")
lines(age.grid, pred$fit-2*pred$se, lty="dashed")
abline(v=c(25,40,60), lty=2, col="darkgreen")
```

The knots are specified at ages 25, 40, and 60. This produces a spline with 6 basis functions (a cubic spline with 3 knots has a 7 (=3+3+1) degrees of freedom).

Alternatively, the `df()` function can be used to produce a spline fit with knots at uniform intervals.

```{r}
dim(bs(Wage$age, knots=c(25,40,60)))
```

```{r}
dim(bs(Wage$age, df=6))
```

```{r}
attr(bs(Wage$age, df=6), "knots")
```

R chooses knots at ages 33.8, 42.0, and 51.0, which correspond to the 25th, 50th, and 75th percentiles of age. The function `bs()` also has a degree argument, so we can fit splines of any degree, rather than the default degree of 3 (which yields a cubic spline).

## Natural Splines 

To fit a natural spline we use `ns()` function. 
```{r}
fit2 <- lm(wage ~ ns(age, df=4), data=Wage)
pred2 <- predict(fit2, newdata=list(age=age.grid), se=T)
plot(Wage$age, Wage$wage, col="grey", 
     xlab = "Age", ylab = "Wage")
lines(age.grid, pred2$fit, col="red",lwd=2) 
```

## Smoothing Splines

We can fit a smoothing spline using `smooth.spline()`
```{r} 
plot(Wage$age, Wage$wage, xlim = agelims, cex = 0.5, col="darkgrey", 
     xlab = "Age", ylab = "Wage")
title(" Smoothing Spline ")
# fit a smoothing spline with degree 16
fit <- smooth.spline(Wage$age, Wage$wage, df = 16)
# chose df using cross-validation
fit2 <- smooth.spline(Wage$age, Wage$wage, cv = TRUE)
# cv yields df=6.8
fit2$df
lines(fit, col="red",lwd=2)
lines(fit2, col="blue",lwd=2)
legend("topright",legend=c("16 DF","6.8 DF"),col=c("red","blue"),lty=1,lwd=2,cex=.8)
```

## A Simulated Example 

Consider the following data set 

```{r}
# simulate data
set.seed(12345)
n  <- 100
x  <- seq(0,1,length=n)
fx <- 5*x*sin(2*pi*x) + 4*cos(3*pi*x)
y  <- rnorm(n,fx,1)
data = data.frame(y,x)
```

Visualize the data 
```{r}
library(ggplot2)
p1 <- ggplot(data,aes(x,y)) +
       geom_point() 
  
p1 + geom_line(aes(x,fx,col="Truth")) +
  scale_colour_manual(" ", values=c("Truth"="black")) 
```

Let's define two knots: 
```{r}
# we have  knots 
k1 <- 0.33
k2 <- 0.62 
p1 + geom_line(aes(x,fx,col="Truth")) +
     scale_colour_manual(" ", values=c("Truth"="black")) +
     geom_vline(xintercept=c(k1,k2), 
                linetype="dotted", color = "blue", size=1)
```


```{r}
data$k1dummy <- ifelse(data$x>k1,1,0)
data$k1diff <- data$x - k1
data$xk1 <- data$k1diff*data$k1dummy
#
data$k2dummy <- ifelse(data$x>k2,1,0)
data$k2diff <- data$x - k2
data$xk2 <- data$k2diff*data$k2dummy
```


```{r}
ggplot(data, aes(x,xk1)) +
  geom_line() +
  geom_line(aes(x,xk2,col="red")) +
  theme(legend.position = "none")
```

Visualization using `bs()` function
```{r}
library(splines)
B = bs(x,knots=c(k1,k2),degre=1)
matplot(x,B,type="l",lty=1,lwd=2)
```

Quadratic spline: 
```{r}
B = bs(x,knots=c(k1,k2),degre=2)
matplot(x,B,type="l",lty=1,lwd=2)
```

Cubic spline: 
```{r}
B = bs(x,knots=c(k1,k2),degre=3)
matplot(x,B,type="l",lty=1,lwd=2)
```

We can easily fit a spline using least squares: 
```{r}
# least squares estimation 
model1  <- lm(y ~ x + xk1 +xk2, data=data)
summary(model1)
yhat <- model1$fitted.values

p1 + geom_line(aes(x,fx,col="Truth")) + 
  geom_line(aes(x,yhat,col="Linear spline")) +
  scale_colour_manual(" ", values=c("Truth"="black","Linear spline"="red")) +
  geom_vline(xintercept=c(k1,k2), 
                linetype="dotted", color = "blue", size=1)
```

<br>
**Question** What kind of fit would you expect if we used a single knot at `k1` and a linear spline? 
<br>

**Here's the answer**
```{r}
# least squares estimation 
model1.a  <- lm(y ~ x + xk1, data=data)
summary(model1.a)
yhat.a <- model1.a$fitted.values

p1 + geom_line(aes(x,fx,col="Truth")) + 
  geom_line(aes(x,yhat.a,col="Linear spline")) +
  scale_colour_manual(" ", values=c("Truth"="black","Linear spline"="red")) 
```


Or using the `splines` library: 
```{r}
library(splines)
model1.bs <- lm(y ~ bs(x, knots = c(k1,k2), degree=1), data=data) 
summary(model1.bs)
yhat.bs <- model1.bs$fitted.values
# compare predictions
plot(yhat, yhat.bs)
sum(yhat-yhat.bs)
```

Note that predictions from the first regression and the one in which we used 
`bs()` function are the same: 
```{r}
sum(yhat-yhat.bs)
```

Let's fit a quadratic spline: 
```{r}
# quadratic fit
model2.bs <- lm(y ~ bs(x, knots = c(k1,k2), degree=2), data=data) 
summary(model2.bs)
yhat2.bs <- model2.bs$fitted.values

p1 + geom_line(aes(x,fx,col="Truth")) + 
     geom_line(aes(x,yhat.bs,col="Linear spline")) +
     geom_line(aes(x,yhat2.bs,col="Quadratic spline")) +
     scale_colour_manual(" ", values=c("Truth"="black", 
                                       "Linear spline"="red", 
                                       "Quadratic spline"="blue")) +
     theme(legend.position = "right")
```

Cubic spline: 
```{r}
# Cubic spline fit
model3.bs <- lm(y ~ bs(x, knots = c(k1,k2), degree=3), data=data) 
summary(model3.bs)
yhat3.bs <- model3.bs$fitted.values

p1 + geom_line(aes(x,fx,col="Truth")) +
     geom_line(aes(x,yhat.bs,col="Linear spline")) +
     geom_line(aes(x,yhat2.bs,col="Quadratic spline")) +
     geom_line(aes(x,yhat3.bs,col="Cubic spline")) +
     scale_colour_manual(" ", values=c("Truth"="black", 
                                       "Linear spline"="red", 
                                       "Quadratic spline"="blue", 
                                       "Cubic spline"="darkgreen")) +
  theme(legend.position = "right")
```

Alternatively, we can set the degrees of freedom (df - model complexity) and let 
the knots be set automatically: 
```{r}
modeldf10  <- lm(y ~ bs(x, 10), data=data)
yhatdf10.bs <- modeldf10$fitted.values
summary(modeldf10)

p1 + geom_line(aes(x,fx,col="Truth")) + 
  geom_line(aes(x,yhatdf10.bs,col="df=10 spline")) +
  scale_colour_manual(" ", values=c("Truth"="black","df=10 spline"="red")) 
```


Fitting a smoothing spline with df=10: 
```{r}
ssfit <- smooth.spline(x, y, df = 10)
p1 + geom_line(aes(x,fx,col="Truth")) +
     geom_line(aes(x,ssfit$y,col="Smooth spline df=10")) +
  scale_colour_manual(" ", values=c("Truth"="black","Smooth spline df=10"="red")) 
```

Choose the df using cross-validation
```{r}
ssfitcv <- smooth.spline(x, y, cv=TRUE)
ssfitcv$df

p1 + geom_line(aes(x,fx,col="Truth")) +
  geom_line(aes(x,ssfit$y,col="Smooth spline df=10")) +
  geom_line(aes(x,ssfitcv$y,col="Smooth spline df=8.85")) +
  scale_colour_manual(" ", values=c("Truth"="black",
                                    "Smooth spline df=10"="red",
                                    "Smooth spline df=8.85"="blue")) 

```


<br>
**Exercise** What happens if you set `df=20` or even `df=30`? Which `df` would you prefer? Why?
<br>

# Local Regression 

In order to perform local regression, we use the `loess()` function.

```{r}
plot(Wage$age, Wage$wage, xlim = agelims, cex = 0.5, col="darkgrey", 
     xlab = "Age", ylab = "Wage")
title("Local Regression")
fit <- loess(wage ~ age, span=.2, data=Wage)
fit2 <- loess(wage ~ age, span=.5, data=Wage)
lines(age.grid, predict(fit, data.frame(age=age.grid)), col="red", lwd=2)
lines(age.grid, predict(fit2, data.frame(age=age.grid)), col="blue", lwd=2)
legend("topright", legend=c("Span=0.2","Span=0.5"), 
       col=c("red","blue"), lty=1, lwd=2, cex=.8)
```

Here we have performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20% or 50% of the observations. The larger the span, the smoother the fit. The `locfit` library can also be used for fitting local regression models in R.

# Generalized Additive Models (GAMs)

A GAM model can be fit using the `lm()` function. For example, to fit a GAM to predict wage using natural spline functions of year and age, treating education as a qualitative predictor, we can use
```{r}
gam1 <- lm(wage ~ ns(year,4) + ns(age,5) + education, data=Wage)
summary(gam1)
```

But the function `lm()` cannot be used to estimate a GAM with smoothing splines. For this we can use `gam` library.

```{r}
library(gam)
gam.m3 <- gam(wage ~ s(year,4) + s(age,5) + education, data=Wage)
```

In the chunk above, the `gam::s()` function is used to indicate that
we would like to use a smoothing spline. The variable year is set to have 4 degrees of freedom (df) and the function of the variable will have 5 df. Because the education variable is qualitative, it is left as is, resulting in the presence of four dummy variables.

The plot function after the GAM estimation is particularly useful to see the shapes of estimated functions: 
```{r}
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE,col="blue")
```

```{r}
plot.Gam(gam1 , se=TRUE , col ="red ")
```

We can perfom a series of `anova()` tests to find the best model. 
In the following

m1: GAM excludes year
m2: GAM uses a linear function of year 
m3: GAM uses a spline function of year (estimated previously) 
```{r}
gam.m1 <- gam(wage ~ s(age,5) + education, data=Wage)
gam.m2 <- gam(wage ~ year + s(age,5) + education, data=Wage)
anova(gam.m1, gam.m2, gam.m3, test="F") 
```

Based on these result, m2 is preferred to m1, that is, a model that includes a linear function of year is better than a model without year. But nonlinear function of year is not needed as we do not reject the null for comparing m2 vs. m3. So we prefer a model with a linear function of year. 

We can use `summary()` function after the GAM estimation to generate a summary of the fitted model.
```{r}
summary(gam.m3)
```

The p-values for year and age correspond to a null hypothesis of a linear
relationship versus the alternative of a non-linear relationship. For the variable year, 
a linear relationship is preferred as indicated by large p-value (0.35). But a nonlinear relationship is required for age. 

Making predictions: 
```{r}
preds <- predict(gam.m2, newdata=Wage)
```

We can use a loess fit as a smoothing term in a GAM formula using the `lo()`
```{r}
gam.lo <- gam(wage ~ s(year, df = 4) + lo(age, span = 0.7) + education, data = Wage)
plot.Gam(gam.lo, se = TRUE, col = "green")
```

In the code chunk above, we used local regression for the age term, with a span of 0.7.
We can also use the `lo()` function to create interactions before calling the
`gam()` function. For example,

```{r}
gam.lo.i <- gam(wage ~ lo(year, age, span = 0.5) + education, data = Wage)
```

We can plot the resulting two-dimensional surface if we first install the `akima` package.
```{r}
library(akima)
plot(gam.lo.i)
```

## Fitting a Logistic Regression using GAM

In order to fit a logistic regression GAM, we use the `I()` function
in constructing the binary response variable, and set `family=binomial`.

```{r}
gam.lr <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = Wage)
par(mfrow = c(1, 3))
plot(gam.lr, se = T, col = "green")
```

Notice the large confidence band in the "<HS" category. This is because there are no high earners in the "<HS" category: 
```{r}
table(Wage$education, I(Wage$wage > 250))
```

We fit a logistic regression GAM using all but this category. This provides more sensible results. Notice how we set the subset option below: 
```{r}
gam.lr.s <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = Wage, subset = (education != "1. < HS Grad"))
plot(gam.lr.s, se = T, col = "green")
```



<br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

